---
title: 'Reproducible Research: Peer Assessment 1'
author: "Shripad Kondra"
date: "28/04/2017"
---

```{r opts, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "figure/"
)
```
## Loading and preprocessing the data

```{r loaddata}
unzip(zipfile="activity.zip")
data <- read.csv("activity.csv")
str(data)
```

## What is mean total number of steps taken per day?
```{r}
#library(ggplot2)
#TotalSteps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
TotalSteps <- aggregate(steps ~ date, data = data, sum)
#qplot(total.steps, binwidth=2000, xlab="Total number of steps taken each day",)
hist(TotalSteps$steps,breaks = 10, main = "Total number of steps taken each day", xlab = "No of Steps ")
```

```{r}
Mn <- floor(mean(TotalSteps$steps, na.rm=TRUE)) # 3
Md <- median(TotalSteps$steps, na.rm=TRUE) # 3
```
Mean of the total number of steps taken per day = `r Mn`

Median of the total number of steps taken per day = `r Md`

## What is the average daily activity pattern?
1. Split data by intervals.
2. Calculate average of steps in each 5 minutes interval.
3. Plot 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
4. Find the interval that contain maximum number of steps. 


```{r}
library(ggplot2)
meansteps <- tapply(data$steps,data$interval,mean,na.rm=TRUE)

plot(row.names(meansteps),meansteps,type="l",
     xlab="Time Intervals (5-minute)", 
     ylab="Mean number of steps taken (all Days)", 
     main="Average Steps Taken at 5 minute Intervals",
     col="red")
```

Find the time interval that contains maximum average number of steps over all days
```{r}
interval_num <- which.max(meansteps)
interval_max_steps <- names(interval_num)
```
The `r  interval_max_steps `th minute  or  `r  interval_num `th 5 minute interval contains the maximum number of steps on average across all the days



## Imputing missing values
1. Calculate and report the total number of missing values in the dataset.
2. Devise a strategy for filling in all of the missing values in the dataset.
3. Create a new dataset that is equal to the original dataset but with the missing data filled
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}
num_na_values <- sum(is.na(data))
```
Total number of missing values in the dataset are `r num_na_values`

Fill in missing values using the **average interval value across all days**
```{r}
na_indices <-  which(is.na(data))
imputed_values <- meansteps[as.character(data[na_indices,3])]
names(imputed_values) <- na_indices
for (i in na_indices) {
    data$steps[i] = imputed_values[as.character(i)]
}
```
Total number of missing values in the dataset after filling are `r sum(is.na(data))`

```{r}
totsteps <- tapply(data$steps, data$date,sum)
hist(totsteps,col="red",xlab="Total Steps per Day", 
      ylab="Frequency", main="Histogram of Total Steps taken per day")
```

New Mean of the total number of steps taken per day = `r ceiling(mean(totsteps))``

New Median of the total number of steps taken per day = `r ceiling(median(totsteps))`



## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
data$date <- as.Date(data$date,"%Y-%m-%d")
days <- weekdays(data$date)

data$day_type <- ifelse(days == "Saturday" | days == "Sunday", "Weekend", "Weekday")
meansteps <- aggregate(data$steps, by=list(data$interval, data$day_type),mean)
names(meansteps) <- c("interval","day_type","steps")
library(lattice)
xyplot(steps~interval | day_type, meansteps,type="l",
       layout=c(1,2),xlab="Interval",ylab = "Number of steps")
```
There is a more activity during week-end time.
